## Example 6
library (ggplot2)
library (gplots)
library (dplyr)

<!--
Set the working directory.
No Rchunk is needed to accomplish this.
-->
Set the working directory.
```{r}
setwd("~/documents/PSY441/CapstoneReproducibility/06 ANOVA")
```
<!--
Read in the homework6.csv data. For each of the following steps you
should preface your R commands with some plain English explanation.
The fabricated data are supposed to approximate how people perform
in different types of visual searches. Specifically, we're 
comparing performance in feature searches (e.g. looking for a red
X among green Xs) and conjunction searches (e.g. looking for a red
X among green Xs and red Os). We're also looking at how the number
of distractors within the search display effect our data.
-->
First we will read in the data. 
```{r}
dat<-read.csv("homework6.csv")
```


<!--
Check the structure of the data and change any variables that should
be factor variables into factor variables
-->
Next we will look at the structure. 
```{r}
str(dat)
```
Change the integer variables so that they will be treated as factor variables. 
```{r}
dat$distractors<-as.factor(dat$distractors)
dat$subject<-as.factor(dat$subject)
dat$subject2<-as.factor(dat$subject2)
```

<!--
Perform a one-way between-subjects ANOVA examining for effects of 
the the number of distractors for the conjunction searches only.
Also create a line graph to help you interpret the results of your
analysis.
-->
A one-way between-subjects ANOVA should be performed for effects of the number of distractors for the conjunction searches. 

The first step is to get only the data from the conjunction searches.

```{r}
temp<-dat[dat$type=="conjunction",]
```

Then we can run our ANOVA.

```{r}
summary(aov(rt~distractors, data=temp))
```

There's a significant effect for the number of distractors in conjunction searches, F(2,21) = 8.6, p < .05. To interpret these data better, I'll create a line graph that shows how the number of distractors affected the performance.

```{r}
library(ggplot2)
library(gplots)
library(dplyr)

temp<-temp%>%group_by(distractors)%>%
  summarize(means=mean(rt),
            sems=sd(rt)/sqrt(length(rt)))

f<-ggplot(temp, aes(x=as.factor(distractors),
                    y=means,
                    group=1))+
  geom_line(size=1)+
  geom_point(size=2)+
  geom_errorbar(aes(ymax=means+sems,
                    ymin=means-sems),
                width=.2)+
  ggtitle("Distraction Effect")+
  labs(x="Distractors",y="Performance")+
  theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
  theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
  theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
  theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
  theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())
f
```

From the figure we can see that for the conjunction searches, as the number of distractions increase, the response time increases as well. 

<!--
Perform a one-way within subjects ANOVA on the same data you analyzed
earlier. Tip: use subject as the subjects variable for this analysis.
-->

I am going to perform a one-way within subjects ANOVA but since I changed temp, the data is not the same as I analyzed last time. In order to perform the appropriate analysis I am going to reuse the code I first used to create temp.

```{r}
temp<-dat[dat$type=="conjunction",]
```

Now I'll perform the within-subjects analysis.

```{r}
summary(aov(rt~distractors+
              Error(subject/distractors),data=temp))
```

The performance of participants decreased as the number of distractions increased, F(2,14) = 56.5, p < .05.

<!--
Perform a between-subjects ANOVA examining for the effects of and
interactions between the number of distractors and the search type.
Make sure to interpret your results.
-->
Now, I'll perform a between-subjects 2-way ANOVA looking for effects
of and interactions between the number of distractors and the search type. 

```{r}
summary(aov(rt~distractors*type, data=dat))
```

The analyses show that there was no effect of distractors, F(2,42) =
2.6, p > .05. There was an effect of type, F(1,42) =
459.4, p < .05. There was an interaction between distractors and type, F(2,42) = 7.8, p < .05.

<!--
Create a line graph and a grouped bar graph to help us interpret
the results of the previous analyses. Include error bars in your
figures.
-->
To help me interpret the earlier analyses, I'll create line and
grouped bar graphs for the data I analyzed. But first I need to
summarize the data.

```{r}
temp<-dat%>%group_by(distractors,type)%>%
  summarize(means=mean(rt),
            sems=sd(rt)/sqrt(length(n)))
```

Now I can create the figures.

```{r}
f<-ggplot(temp, aes(x=distractors,
                    y=means,
                    group=type,
                    color=type))+
  geom_line(size=1)+
  geom_point(size=2)+
  scale_color_manual(values=c("darkgreen","red"),
                     name="Type",
                     breaks=c("no","yes"),
                     labels=c("No", "Yes"))+
  geom_errorbar(aes(ymax=means+sems, ymin=means-sems),width=.2)+
  ggtitle("Performance and Distraction by Search Type")+
  labs(x="Distractors",y="Response Time")+
  theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
  theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
  theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
  theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
  theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())+
  theme(legend.position=c(.4,.76))+
  theme(legend.background=element_blank())+
  theme(legend.background=element_rect(color="black"))+
  theme(legend.title=element_blank())+
  theme(legend.title=element_text(size=12))+
  theme(legend.title.align=.5)+
  theme(legend.text=element_text(size=10,face="bold"))
f

f<-ggplot(temp, aes(x=distractors,
                    y=means,
                    fill=type))+
  geom_bar(stat="identity",position=position_dodge())+
  scale_fill_manual(values=c("darkgreen","red"),
                     name="Type",
                     breaks=c("no","yes"),
                     labels=c("No", "Yes"))+
  geom_errorbar(aes(ymax=means+sems,
                    ymin=means-sems),
                width=.2,
                position=position_dodge(.9))+
  ggtitle("Performance and Distractions by Search Type")+
  labs(x="Distractors",y="Response Time")+
  theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
  theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
  theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
  theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
  theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())+
  theme(legend.position=c(.4,.76))+
  theme(legend.background=element_blank())+
  theme(legend.background=element_rect(color="black"))+
  theme(legend.title=element_blank())+
  theme(legend.title=element_text(size=12))+
  theme(legend.title.align=.5)+
  theme(legend.text=element_text(size=10,face="bold"))
f
```
<!--
Interpret the statistical results in light of the figures.
-->

From the figures and earlier analyses we can see that the response time increases as the number of distractions increase. 

<!--
Perform a within-subjects two-way ANOVA examining for the effects
of and interactions between number of distractors and search type.
Tip: use subject as the subject's variable for this analysis.
Interpret the results of the analysis based on the statistical
results and your earlier figures.
-->

Here I'm performing a within-subjects two-way ANOVA to examine for
the effects of and interactions between the number of distractors and search type.

```{r}
summary(aov(rt~distractors*type+
              Error(subject/(distractors*type)),
                    data=dat))
```

From the analyses and the figures we can see that there is a difference in response time in response to the number of distractions, F(2,14) 11.2, p < .05. We can also see that there is a difference in response times depending on the type of search, F(1,7) = 1626, p < .05. Finally,we can see that people perform differently in response to the distractors depending on which type of search they perform, F(2,14) = 32.1, p < .05.

<!--
Finally conduct and interpret the results of a mixed ANOVA in which
number of distractors varies within subjects and the type of search
varies between subjects. Make sure to interpret your results. Tip:
use subject2 as your subject's variable for this analysis.
-->

```{r}
summary(aov(rt~distractors*type+
              Error(subject2/distractors),
            data=dat))
```
From our analyses, we can see that individuals performed differently depending on the type of search they performed F(1,14) = 182.4, p < .05. Additionally we see that performance was different depending on the number of distractions F(2,28) = 11, p < .05. Finally, we see that in feature searches the performance of participants 
increased or stayed the same with increased distractions and on the conjunction searches the performance of participants decreased as the number of distractions increased, F(2,28) = 32.5, p < .05.
