## Homework 5

<!--
Set the working directory.
No Rchunk is needed to accomplish this.
-->
Set the location of the working files.
```{r}
setwd("~/documents/PSY441/CapstoneReproducibility/05 t-tests")
```

<!--
Read in the homework5.csv data. For each of the following steps you
should preface your R commands with some plain English explanation.
The data are fabricated and include the number of words per minute
that people can type when listening to no music and when listening
to heavy metal.
-->
Load in the data.
```{r}
dat<-read.csv("homework5.csv")
```


<!--
Check the structure of the data.
-->

Look at the structure of the data. 
```{r}
str(dat)
```
We can see that dat is a data frame with two variables of music: metal and none.
<!--
Find the mean words per minute for the two groups.
-->
Find the mean words per minute for the two groups. 

```{r}
mean(dat$wpm[dat$music=="metal"])
mean(dat$wpm[dat$music=="none"])
```

By looking at the means for both groups it appears that the metal group typed more words per minute than the no music group.
<!--
Run an independent t-test to determine if the mean wpm for the no music group is different from the mean wpm from heavy metal group. Interpret the results of your analysis after you run it.
-->


I will run an independent t-test to determine if the mean wpm for the
metal music group is different from the mean wpm of the no music group. 
```{r}
t.test(dat$wpm[dat$music=="metal"],
       dat$wpm[dat$music=="none"],
       var.equal=TRUE)
```

The results of the t-test show that the mean words per minute for the metal music group is different from the mean words per minute of the no music group, t(62) = 2.33, p = .0230.
<!--
Run another independent t-test with the alternative hypothesis that
the heavy metal group types faster than the no music group. Interpret
the results of your analyses.
-->
I will run an independent t-test with the alternative hypothesis
that the metal music group typed faster than the no music group.

```{r}
t.test(dat$wpm[dat$music=="metal"],
       dat$wpm[dat$music=="none"],
       var.equal=TRUE,
       alternative="greater")
```
The results of the t-test show that participants in the metal music group typed faster than participants in the no music group, t(62) = 2.33, p = .0115.

<!--
Create a bar graph of the data.
-->
To create a figure of the data I need to load the
ggplot2, gplots, and dplyr libraries.

```{r}
library(ggplot2)
library(gplots)
library(dplyr)
```

Next I need to use dplyr to summarize the data and find the means and
standard errors of the means for the different groups.
```{r}
temp<-dat%>%group_by(music)%>%
  summarize(means=mean(wpm),
            sems=sd(wpm)/sqrt(length(wpm)))
```
Now that the data is summarized I can create a figure of the data.

```{r}
f<-ggplot(temp,aes(x=factor(music),y=means))+
  geom_bar(stat="identity", color="black",
           fill=c("deepskyblue2", "deeppink"))+
  geom_errorbar(aes(ymax=means+sems,
                    ymin=means-sems),
                width=.2)+
  ggtitle("Typing Speed For the Groups")+
  labs(x="Music", y="words per minute")+
  scale_x_discrete(breaks=c("none","metal"),
                   labels=c("none","metal"))+
  theme(plot.title=element_text(size=15,
                                face="bold",
                                vjust=.5))+
  theme(axis.title.x=element_text(size=12,
                                  face="bold",
                                  vjust=-.25))+
  theme(axis.title.y=element_text(size=12,
                                  face="bold",
                                  vjust=.25))+
  theme(axis.text.x=element_text(size=10,
                                 face="bold",
                                 color="black"))+
  theme(axis.text.y=element_text(size=10,
                                 face="bold",
                                 color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),
        axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())
f
```
<!--
Conduct and interpret a two-tailed paired t-test on the data.
-->


Now I will conduct and interpret a two-tailed paired t-test on the data.
```{r}
t.test(dat$wpm[dat$music=="none"],
       dat$wpm[dat$music=="metal"],
       var.equal=TRUE,
       paired=TRUE)
```
Results show that the typing speed was different for the two groups. The means indicate that the participants in the metal music group typed faster than the participants in the no music group, t(31) = -12.153, p < .05.